---
title: "RepData_PeerAssessment1"
author: "Christophe Cahn"
date: "23 janvier 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading libraries

```{r library loading, echo=TRUE}
library(dplyr)
library(data.table)
library(ggplot2)
```

## Loading and preprocessing the data

```{r loading data, echo=TRUE}
data <- read.csv("./data/activity.csv", header = TRUE, sep = ",")
data$date <- as.Date(data$date)
data <- tbl_df(data)
```

## What is mean total number of steps taken per day?

We now compute the number of steps by day, using the 'summarise()' function.
```{r grouping per day, echo=TRUE}
daily <- group_by(data, date)
(per_day   <- summarise(daily, steps = sum(steps,na.rm = TRUE)))
```

```{r histogram 1, echo=TRUE}
hist(per_day$steps, breaks = 15)
rug(per_day$steps)
```

```{r, echo=TRUE}
mean(per_day$steps, na.rm = TRUE)
median(per_day$steps, na.rm = TRUE)
```

## What is the average daily activity pattern?

We first need to compute average over 5-minute interval.
```{r, echo=TRUE}
inter <- group_by(data, interval)
(per_interval <- summarise(inter, steps = mean(steps,na.rm = TRUE)))
plot(per_interval$interval,per_interval$steps, type = "l")
nb.step.max=max(per_interval$steps,na.rm=TRUE)
filter(per_interval,steps==nb.step.max)$interval
```

Seems to be the 835 interval.

## Imputing missing values

```{r, echo=TRUE}
dim(filter(data,is.na(steps)))
```

Then, we find 2,304 NA values.

To fill the NA, we'll use the average value for the correpsonding 5-minute interval. We use the merge function for this purpose.

```{r, echo=TRUE}
m <- merge(data,per_interval,by="interval")
m2 <- mutate(m,steps=steps.x)
m2 <- tbl_df(m2)
m2$steps[is.na(m2$steps.x)] <- m2$steps.y[is.na(m2$steps.x)]
newdata <- arrange(select(m2,steps,date,interval),date)
```

The dataset 'newdata' is then equal to the original dataset but with the missing data filled in.

```{r, echo=TRUE}
by_day <- group_by(newdata, date)
steps_per_day <- summarise(by_day,steps = sum(steps))
```

```{r, echo=TRUE}
hist(steps_per_day$steps, breaks = 15)
rug(steps_per_day$steps)
```

```{r, echo=TRUE}
mean(steps_per_day$steps)
median(steps_per_day$steps)
```

The cleaning operation has an impact on the mean and median of the new distribution.

## Are there differences in activity patterns between weekdays and weekends?

Let's identify the weekdays.

```{r weekdays identification, echo=TRUE}
Sys.setlocale("LC_TIME", "en_US.UTF-8")
kub <- mutate(newdata,day=weekdays(date))
kub$day[kub$day=="Saturday" | kub$day=="Sunday"] <- "weekend" 
kub$day[kub$day!="weekend"] <- "weekday" 
kub$day <- factor(kub$day)
table(kub$day)
```

We then compute average for weekdays and weekends.
```{r computing average over factors, echo=TRUE}
step_weekdays <- summarise(group_by(filter(kub,day=="weekday"),interval), steps = mean(steps))
step_weekends <- summarise(group_by(filter(kub,day=="weekend"),interval), steps = mean(steps))
```
Now the plots :
```{r plotting panel, echo=TRUE}
par(mfrow=c(2,1))
plot(step_weekdays, type = "l", main = "Weekdays", col="blue")
plot(step_weekends,type = "l", main = "Weekends", col="red")
```
